Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-10-10 doc/paper-lcr.tex
appendix on kappa inference added

@@

%\section{Smooth component models}
%\label{appendix:smooth}

>>>>

\todo{Matt}{Edit the above text to add a note on how we do the sampling in
practice, via the CDF.}

@@

\section{Accounting for uncatalogued low mass halos, and voids}
\label{appendix:smooth}

Our halo model allows us to infer a halo mass $\Mhalo$ for each galaxy in a
photometric catalog; under the weak lensing approximation described in
\Sref{sec:theory} we can compute the contribution of each of these halos to
the overall convergence and shear amplitude, assuming a homogeneous
Friedman-Robertson-Walker metric and the concordance $\Lambda$CDM cosmological
parameters for the angular diameter distances to calculate the critical
density for each lens plane. Clearly this approach will tend to over-estimate
the convergence, since we are placing massive halos into a volume that has
already been assumed to be full of homogeneously distributed matter with the 
mean cosmological density. In practice, the space between massive halos will
contain a) low mass halos that are not bright enough to be detected and b)
empty space where the density is below that of the mean. A rigorous treatment
of this complex situation is beyond the scope of this paper, but is being
investigated (Blandford et al, in preparation). In this appendix we describe
our empirical approach to accounting for the missing mass and voids.

Let the summed convergence due to halos in our galaxy catalog be $\kappah$.
Given some assumptions about halo density profile and shape, this parameter
can be inferred from a set of uncertain halo mass estimates (which have in
turn been inferred from noisy measurements of galaxy redshift and stellar mass
as described in \Aref{appendix:MSMH}). We assume each halo to be a
spherically-symmetric NFW profile, and neglect the contribution of the stellar
mass to the convergence. The halo mass inference for the $k^{\rm th}$ galaxy 
is stored as a list of sample values of $\Mhalok$ drawn from the posterior PDF
$\Pr(\Mhalok|\Mstarobsk,z_k)$; computing the contribution to the convergence
due to this halo at the target line of sight for each sample gives, in turn, 
a set of
samples drawn from the PDF $\Pr(\kappahk|\Mstarobsk,z_k)$. We then generate
samples from $\Pr(\kappah|\{\Mstarobsk,z_k\})$ by drawing a sample $\kappahk$
for each halo, summing over halos to get $\kappah$, discarding those
samples and moving down the lists.

The PDF $\Pr(\kappah|\{\Mstarobsk,z_k\})$ will, in general, be broad (due to
the uncertainties involved in halo mass estimation). It will also be shifted
towards high values of convergence, due to the FRW approximation described
above. What we really want is an inference of $\Pr(\kappa|\{\Mstarobsk,z_k\})$
instead. We can obtain this by considering the expansion:
\begin{equation}
\Pr(\kappa|\{\Mstarobsk,z_k\}) = \int d\kappah 
   \Pr(\kappa|\kappah) \Pr(\kappah|\{\Mstarobsk,z_k\})
\label{eq:kappaconv}   
\end{equation}
The first term in the integrand relates the summed  convergence due to model
halos, $\kappah$, to the true summed convergence, $\kappa$.  In the Millenium
Simulation catalogs, we can compute a single value of $\kappah$ for each
selected line of sight from the true halo mass and redshift (and our same
assumptions of halo profile and shape), and tabulate the two-dimensional PDF
$\Pr(\kappa|\kappah)$ as a sequence of one-dimensional PDFs for $\kappa$ in a
bin at fixed $\kappah$. Note that this PDF captures the ``intrinsic scatter''
of $\kappah$ due to our assumptions about the clumpiness of mass in the
Universe; the integral is the convolution of this distribution with that
arising from our observational uncertainties.

For any given observed catalog then, we infer
$\Pr(\kappah|\{\Mstarobsk,z_k\})$ as described above, and then
multiply it by $\Pr(\kappa|\kappah)$; we
again do the marginalisation over $\kappah$ by Monte Carlo, drawing samples
from the two-dimensional grid and keeping only the $\kappa$ values, to form
our final result, $\Pr(\kappa|\{\Mstarobsk,z_k\})$. 
This PDF will describe, by construction, an accurate (i.e., unbiased)
measurement of $\kappa$ in the case where the Millenium Simulation halo
catalog and Behroozi et al $\Mstar-\Mhalo$ relation are themselves accurate
descriptions of galaxies and their distribution. In \Sref{sec:tests} we
investigate the amplitude of the systematic errors incurred if these
assumptions are not valid.

\todo{Matt}{Add text describing whatever sampling shortcut you derive! :-)}

% Finally, we note that \Eref{eq:kappaconv} and the method for evaluating the
% integral in it can be generalized to yield 
% $\Pr(\kappa,|\gamma||\{\Mstarobsk,z_k\})$, where $|\gamma|$ is the amplitude
% of the complex shear. The amplitude of the shear due to halos, $|\gammah|$ is
% needed in this calculation, and marginalised over -- $\gammah$ is computed by
% simple summation of its components, as in \Sref{sec:theory}. In
% \Sref{sec:includinggamma} we explore the impact that knowing $\gamma$ (from,
% for example, a strong lens model)  has on the inference of $kappa$, and how
% severe the systematic errors from getting this wrong might be.
